---
title: "Homework_Week2"
author: "Meaghan Burke"
date: "September 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2)
library(ggplot2)
```

### Week 2 Assignment 

**HW - HA (6.2)**


2. The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years. 



a. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?

**Trend: Increasing over 1 year**
**Appears to be some seasonality(lag graph), length 1 year** 

```{r}

autoplot(plastics) + ggtitle("Sales of Product A") + ylab("Sales(Monthly)")
ggAcf(plastics)+ ggtitle("Seasonality of Product A") 
```


b. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

** The multiplicative decomposition shows a general upward trend from year 2 through year 5. After year 5, the trend appears to be decreasing. The decomposition graph also displays a seasonal shift (up/down) in one year increments.**
```{r}
plastics %>%
decompose(type="multiplicative") %>% 
autoplot() + 
ggtitle("Sales of Product A")
```




c. Do the results support the graphical interpretation from part a?

** Both part a and part b display a seasonal component of 1 year and a trend that is increasing over time. The decomposition does show a decline after year 5 because it relies on moving averages.**
d. Compute and plot the seasonally adjusted data.

**The "Adjusted Seasonal" plot shows the sales with the seasonality removed, making the upward trend is more visable**
```{r}

mult.decomposition <- plastics %>%
  decompose(type="multiplicative") %>%
  seasonal


autoplot(plastics /mult.decomposition, series = "Adjusted Seasonal")+ 
  autolayer(plastics, series = "Plastics - Normal") +
  ggtitle("Sales of Product A & Seasonally Adjusted Sales Product A")
```


e. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

**After 500 was added to the 35th datapoint in "plastics.adj" dataset, a large spike occurred in the middle of the data series and the seasonally adjusted data series. This can be observed in the chart below.**


```{r}
plastics.adj <- plastics
plastics.adj[35] <- plastics.adj[35]+500


mult.decomposition.adj <- plastics.adj %>%
  decompose(type="multiplicative") %>%
  seasonal


autoplot(plastics.adj /mult.decomposition, series = "Adjusted Seasonal")+ 
  autolayer(plastics.adj, series = "Plastics - Normal") +
  ggtitle("Sales of Product A & Seasonally Adjusted Sales Product A (Adjusted Series [MIDDLE])")
```

f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?



**After 500 was added to the 60th datapoint in "plastics.adj.end" dataset, a large spike occurred at the end of the dataset. However, the seasonality seems to more closely mirror that of the original chart. ** 


```{r}
end.series<- length(plastics)
plastics.adj.end <- plastics
plastics.adj.end[end.series] <- plastics.adj.end[end.series]+500


mult.decomposition.adj <- plastics.adj.end %>%
  decompose(type="multiplicative") %>%
  seasonal


autoplot(plastics.adj.end /mult.decomposition, series = "Adjusted Seasonal [END]")+ 
  autolayer(plastics.adj.end, series = "Plastics - Normal") +
  ggtitle("Sales of Product A & Seasonally Adjusted Sales Product A (Adjusted Series [END])")
```
