---
title: 'Data 624: Week 4 Homework'
author: "Angrand, Burke, Deboch, Groysman, Karr"
date: "October 12, 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
### Week 4 Assignment 


**Chapter 7 HA 7.1, 7.3**


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2)

```

#####  7.1 Consider the **pigs** series - the number of pigs slaughtered in Victoria each month.
The pigs ts contains monthly total number of pigs slaughtered in Victoria, Australia (Jan 1980 to Aug 1995).

```{r, eval=TRUE}
#sampling and shape of dataset, preliminary EDA
head(pigs)

#autoplot the series
pigsdata<-window(pigs)
autoplot(pigsdata) +
  ylab("Number of pigs slaughtered") +xlab("Year")
```




  **a.** Use the **ses()** function in R to find the optimal values of ${\alpha}$ and ${\ell}_{0}$, and generate forecasts for the next four months.
  
```{r, eval=TRUE}
#Estimate parameters
fc_pigs_ses<-ses(pigsdata, h=4)



#Get forecasted estimate parameters from model -(alpha and l)
round(fc_pigs_ses$model$par[1:2],4)

#generate 4 months of forecasts 
data.frame(fc_pigs_ses)

#Accuracy of one-step-ahead training errors
round(accuracy(fc_pigs_ses),2)

# see how SES model was fitted
fc_pigs_ses$model

# get 1st 4 months of forecasts
tsCV(pigs,ses,h=4)[1:4,]

#Plot (Note that forecast using ses doesn't have a trend component.)
fc_pigs_ses<-ses(pigs, h=100)
autoplot(fc_pigs_ses) + 
    autolayer(fitted(fc_pigs_ses), series="fitted") +
    ylab("Total number of Pigs slaughtered in Victoria, Austrailia") + xlab("Jan 1980 to Aug 1995")
```


  **b.** Compute a 95% prediction interval for the first forecast using $\hat{y} {\pm} 1.96s$
  is the standard deviation of the residuals. Compare your interval with the interval produced by R.

```{r, problem_7_1, eval=TRUE, include=TRUE}
# 95% prediction interval for the first forecast

lower.upper <- data.frame( fc_pigs_ses$lower[1, "95%"], fc_pigs_ses$upper[1, "95%"])
names(lower.upper)<- c("lower.limit", "upper.limit")
lower.upper


# calculate standard deviation of residuals with and without model, s = 10273.69 vs s (estimated) 10308.58
s <- sd(fc_pigs_ses$residuals)
print(paste("Standard Deviation: ",round(s,2)))


# calculate 95% prediction interval with model
pred.interval.model<- data.frame(fc_pigs_ses$mean[1] - 1.96*s,fc_pigs_ses$mean[1] + 1.96*s)
names(pred.interval.model)<- c("Lower.Model", "Upper.Model")
pred.interval.model

# calculate 95% prediction interval without model
pred.interval.womodel<- data.frame(mean(pigs)  - 1.96*s, mean(pigs) + 1.96*s)
names(pred.interval.womodel)<- c("Lower.NOModel", "Upper.NOModel")
pred.interval.womodel
```

  - R gives an interval of [78611.97, 119020.8] and by computing the standard deviation of the residuals we got [78679.97, 118952.8].  Both are really close.




#####  7.3 Modify your function from the previous exercise to return the sum of squared errors rather than the forecast of the next observation. Then use the **optim()** function to find the optimal values of ${\alpha}$ and ${\ell}_{0}$. Do you get the same values as the **ses()** function?

  - my_ses_func returns an l value that is ~.01% different then the ses function.

```{r echo=FALSE, message=FALSE, warning=FALSE, Question3}
# modify ses function to return SSE
my_ses_func <- function(pars=c(alpha, level), y){
  Err <- 0
  SSE <- 0 # calculation varies from 
  for(i in 1:length(y)){
    Err <- y[i] - pars[2]
    SSE <- SSE + Err^2
    pars[2] <- pars[1] * y[i] + (1 - pars[1]) * pars[2]
  }
  return(SSE)
}

# compare ses and my_ses_fun output using the 
opt_ses_pigs <- optim(par = c(alpha = 0.2971, l = pigs[1]), y = pigs, fn = my_ses_func)

round(opt_ses_pigs$par,4)
round(fc_pigs_ses$model$par,4)
```
  
