---
title: 'Data 624: Project 1: Part B'
author: "Angrand, Burke, Deboch, Groysman, Karr"
date: "October 19, 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
fig_width: 7
fig_height: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(xlsx)
library(ggplot2)
library(gridExtra)
```


### Part B - Forecasting Power

**Data: ResidentialCustomerForecastLoad-624.xlsx**

Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.  Your assignment is to model these data and a monthly forecast for 2014. The data is given in a single file.  The variable 'KWH' is power consumption in Kilowatt hours, the rest is straight forward. Add these to your existing files above - clearly labeled.  


  **a.** Read in File 
```{r message = F, warning = F}
temp = tempfile(fileext = ".xlsx")
dataURL <- "https://raw.githubusercontent.com/mburke65/CUNY_Data624/master/ProjectFolder/Provided_Files/ResidentialCustomerForecastLoad-624.xlsx"
download.file(dataURL, destfile=temp, mode='wb')

power.data <- readxl::read_excel(temp, sheet =1)


```
  **b.** EDA Analysis 
    - Check/Fill in null values 
    - Convert to time series
    - Graph the monthly data 
      - General plot & seasonal plot: seasonality can be observed in the below plot. There are spikes each year from May to August (air conditioning?) and again in December (holiday season?). There is a slight dip in Jul 2010 maybe due to an unseasonably cold month.
      - Seasonal Box Plot: provides a similar visual to the seasonal plot with usage spikes in the summer months and December. IT also highlights the flucuations in consumption within each month. 
      - Decomposition components graph: this plot again shows that there is a general upwards trend in the data with an observed outlier in July 2010.  
```{r}
head(power.data)
summary(power.data)
print(paste("Check for nulls: ",sum(is.na(power.data)), " Row of Nulls"))
power.data[is.na(power.data)] <- median(power.data$KWH,na.rm = TRUE)
power.data.ts <-ts(power.data[,"KWH"],start = c(1998,1),frequency = 12)
ggtsdisplay(power.data.ts, points = FALSE, plot.type = "histogram") 
ggseasonplot(power.data.ts)
boxplot(power.data.ts~cycle(power.data.ts),xlab="Month", ylab = "Power Usage")
plot(decompose(power.data.ts))


```
  **C.** Model 1: Arima W/ Box-Cox Transformation
    - Replace outlier with tsoutlier suggestion (utilizes a box-cox transformation)
    - Use an auto arima model on the box-cox adjusted data
      - Suggested model: ARIMA(0,0,3)(2,1,0)[12] with drift. RSME(595389) & AICc (5332.67)
    - Check the residuals to make sure the model is satisfactory: 
      - ACF /PACF Plots: the residual appears normal residuals mostly around 0, suggesting stationarity of the residuals 
      - The Box Ljung tests presents a p-value of 0.6951 which indicates white noise
    - Forecast 2014 power values & plot forecasted values  


```{r}
#outlier detection/suggestion/replacement
find.outlier<- tsoutliers(power.data.ts, iterate = 2, lambda = "auto")
power.data.ts.bc<- power.data.ts
power.data.ts.bc[find.outlier$`index`[1]] <- find.outlier$replacements[1]
print(paste("Suggested/Implemented Change for Outlier: ",power.data.ts.bc[151], " Original Value",power.data.ts[151]))

#auto arima model
power.model <- auto.arima(power.data.ts.bc, seasonal = TRUE, stepwise = FALSE)
summary(power.model)
#check residuals
checkresiduals(power.model)

#forecast model @ 95%
forecast.power <- forecast(power.model, level = c(95), h =12)
forecast.power
autoplot(forecast.power)

```

  **D.** Model 3: ETS W/ Box-Cox Transformation
    - Utilize the transformed data & ETS model
    - Model Results: RMSE (630869.7) & AICc (6148.032)
    - Check the residuals to make sure the model is satisfactory: 
      - ACF /PACF Plots: the residual appears normal residuals mostly around 0, suggesting stationarity of the residuals 
      - The Box Ljung tests presents a p-value of 0.0002921 which may indicate that there's dependency issues with the lags
      - 
```{r}
#model w/ previously transformed data
power.model.ets <- ets(power.data.ts.bc)
summary(power.model.ets)

#check residuals
checkresiduals(power.model.ets)

#forecast model @ 95%
forecast.power.ets <- forecast(power.model.ets, level = c(95), h =12)
forecast.power.ets
autoplot(forecast.power.ets)


```
    


  **E.** Model 3: STL 
    - STL model will be the secondary model as it provides the user more control and can be robust when dealing with outliers.
    - Model summary: RMSE (843202.3) & AICc (6261.559)
    - Check residuals: 
      - ACF/PACF: most lags are within the error bounds, suggesting stationarity of the residuals 
      - Box Ljung:p-value = 0.1032 which indicates white noise
    - Forecast 2014 power values & plot forecasted values  
```{r}
power.model.stl <- stlf(power.data.ts, damped=TRUE,  s.window='periodic', robust=TRUE) 
summary(power.model.stl)

power.model.stl<- forecast(power.model.stl)
autoplot(power.model.stl, 12)

checkresiduals(power.model.stl)
```


